---
title: "Trade_SSA"
output: html_document
---


#Loading relevant libraries
```{r message=FALSE, warning=FALSE}
library(tidyr)
library(reshape2)
library(ggplot2)
library(readxl)
library(tidyverse)  # using tidyverse (or dplyr) will enable using filter()  and rename()
```


#Load excel using readxl library
```{r}
Trade_SSA <- read_excel("Trade_SSA.xlsx")
View(Trade_SSA)
```


#Exploring data
```{r}
str(Trade_SSA)
Trade_SSA <- Trade_SSA[-c(4)]
ls(Trade_SSA)
```





# Coverting the data from Wide format to long format using reshape2 library
```{r}
Trade_SSA <- melt(Trade_SSA, id.vars = c("Reporter Name", "Partner Name", "Trade Flow", "Indicator"), measure.vars = c("1992", "1993","1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016"), variable.name = "Years", value.name = "Trade_Value")
```


#Renaming column
``` r 
colnames(Trade_SSA)[colnames(Trade_SSA)=='Partner Name'] <- 'Partner'
colnames(Trade_SSA)[colnames(Trade_SSA)=='Reporter Name'] <- 'Country'
colnames(Trade_SSA)[colnames(Trade_SSA)=='Trade Flow'] <- 'Trade_Flow'
```

```{r}
Trade_SSA <- Trade_SSA %>% 
  dplyr::rename(Trade_Flow = `Trade Flow`) %>% 
  dplyr::rename(Country = "Reporter Name")
```



#line chart
```{r, fig.width=12, fig.height=6}
p1 <- ggplot(data=Trade_SSA, mapping = aes(x=Years, y= Trade_Value))
p1 +geom_line(aes(group=Country,  color= Country), size= 1) + facet_grid(~ Trade_Flow) + theme(axis.text.x=element_text(angle=-55, hjust=0), legend.position = "top")
#ggsave("trade.pdf")
```

## Reactivity is an issue.  

The workshop was recorded.  https://library.capture.duke.edu/Panopto/Pages/Viewer.aspx?id=7a59e23a-1f7f-4bd7-8ebc-a943014170b4 

Check Mine's discusion of reactivity. From timestamep 1:42:40 till about 2:00:00

Reactivity can be used to filter by country.  


Mine was dealing with a factor data type so she used the `levels()` function.  In this case, `Trade_SSA$Country` is a characer data type.  The equivalent function, then, is `unique()`.  See this in action in the ggplot code chunk below.


```{r}
unique(Trade_SSA$Country)
```


Below is an example of how the line graph will displayed if it is filtered by a value.  Eventually, via Shiny, the user will select the country which will effect the filtering of the data frame.  

```{r ggplot}
Trade_SSA %>% 
  filter(Country %in% c("Germany", "India")) %>% 
  ggplot(aes(Years, Trade_Value)) +
  geom_line(aes(group = Country, color = Country), size = 1) +
  facet_grid(~ Trade_Flow) + 
  theme(axis.text.x=element_text(angle=-55, hjust=0), legend.position = "top")
```
